Writing Trash Bin

### 2024-06-20 writingdumpster.qmd
a large amount of restructuring

@@

>>>>

For example, @Hofman2020 asked participants to judge the effectiveness of a particular treatment. One group was shown the prediction interval around the mean, which indicates the uncertainty associated with treatment *for a particular person* and another group was shown the the sampling interval which depicted the uncertainty associated with the *average effectiveness of the treatment*.  The contribution of this experiment hinges on the belief that sharing a common descriptive statistic (in this case, the mean) is grounds to compare two unrelated inferential statistics, a belief which is clearly false when viewing uncertainty through the lens of inference.

@Boukhelifa2012 tried to quantify the strength of the intuitive connection between a line attribute called "sketchiness" and uncertainty. Participants were shown the six scenarios depicted in @fig-sketchy and asked to interpret what they believed the squiggly line indicates. The authors were aware that uncertainty seemed to have some "task" dependence but interpreted this as *context* dependence rather than dependence on a particular inferential statistic. Therefore, "sketchiness" was added to aspects of the visualization that had no obvious inferential statistic associated with it, so participants simply ignored the "sketchiness" or assumed it represented something else, such as an alternative option.

![The graphics displayed by @Boukhelifa2012 to identify if there is an intuitive connection between sketchiness and uncertainty. These graphic were made without concern as to what (i.e. which statistic) in the image is supposed to be uncertain. This leads to the images being difficult to connect to uncertainty even if we assume tha that is what it is supposed to represent. For example, (S5) implies that participants should have read the plot to mean that the creators of a rail network map would be "uncertain" about the existence of a train line, something that is defies common logic.](sketchiness.png){#fig-sketchy width=50%}

@Padilla2021 found that high uncertainty in the model estimates and low forecaster confidence  both caused participants to have decreased confidence in their results and suggested modelers express both if they are relevant.

@Kale2019 discusses the importance of communicating decisions made in the data analysis pipeline and being aware of the alternatives.

@@

# Great Examples 
Despite the common problems detailed in the previous sections, there is some interesting work in the uncertainty visualisation space. This can come in the form of an uncertainty visualisation that attempts to visualise a typically ignored aspect of uncertainty, or an experiment that avoids the pitfalls detailed in the previous sections.

The literature focus on "quantifiable" uncertainty leaves the variance that occurs at early stages in our analysis that is difficult to quantify ignored and forgotten. Some authors have chosen to focus on these unquantifiable cases and look at expressing more complicated cases of uncertainty.  @Tierney2023 builds upon the tidy data principles to allow users to handle missing values. This includes data plots with a missing value "shadow" that allows visualisation authors to identify if the variables used in a plot have any structure in their missing values, which would contribute to uncertainty. Another example of uncertainty that is often ignored is the uncertainty resulting from human choices. Climate scenario uncertainty, shown in @fig-climatescenario, attempts to display the range of climate change outcomes that can result from a range of best and worse case human choices.

![Notice: I will make my own Eversion of a climate scenario plot, this one is just here from a random site as a place holder. A climate scenario plot shows the forecasts of different emission scenarios depending on human choices. The model visualisation wants to show the impact of human choice on the visualisation, rather than assume it away, so five different cases are shown in the model. Each of the case studies also depicts the statistical uncertainty of the model using a confidence interval. ](climatescenario.png){#fig-climatescenario}

Another visualisation method that has a lot of potential is visualisation of samples. Visualisations that opt to express a signal as a sample rather than an estimate have the potential to suppress signal since it is not explicitly visualised, however this has yet to be shown in evaluation experiments. This is not to say that visualisations of mass would not be able to perform signal suppression, but a sample can easily be expressed using aesthetics such as colour on a map and mass visualisation often struggling with issues such as over or under smoothing. These sampling methods can show more of the messiness of the data that sits behind a model. This may not have a detrimental effect on the viewers ability to extract global statistics, as it seems they can be extracted from a visualisation of a sample with ease [Franconeri2021]. Sample visualisations have been used in maps with the pixel map, shown in @fig-samples [Lucchesi2017], but is more commonly used in animation with the HOPs plots [@Hullman2015] or similar concepts [@Blenkinsop2000]. This method has also been adopted by visualisation authors outside of academia as can be seen in the  and the New York Times class mobility figure, shown in @fig-samples, an animated version of which can be found [here](https://www.nytimes.com/interactive/2018/03/19/upshot/race-class-white-and-black-men.html).

::: {#fig-samples layout-ncol=2}

![Pixel map](pixelmap.png){#fig-pixelmap}

![Outcome plot](classmobility.png){#fig-nyt}

Notice: Will replace both plots with my own r visualisations. The pixel map depicts a map of the different poverty rates in local government areas in the US state of Missouri. The NYT visualisation shows an screencap of an animation that depicts the economic status of white and black boys who were raised in wealthy families. By displaying a psudo-sample rather than an estimate and a variance, the visualisation does some signal suppression as the true poverty rate is masked and can only be extracted by taking the global statistic of a sample. You can also depict the signal using explicit values, as the the NYT visualisation did, but it is important that the main visual features depict the uncertainty in the estimate.
:::

A final method to perform signal suppression is simply to visualise the data if it is available and relevant to the uncertainty distribution. An example of this is shown in @fig-census for racial distributions spatially in America, an interactive version of the plot can be found [here](https://edition.cnn.com/interactive/2021/us/census-race-ethnicity-map/). This map shows the typical causes of uncertainty in a spatial model, (e.g. regions where data is sparse, ethnically diverse areas, uneven distribution of points within boundaries, etc) but it avoids the need to create a visualisation with a specific signal in mind. This is the technique typically employed by exploratory data analysis, which means it's lack of a specified signal means there is both *no* uncertainty (since we are technically not performing inference) but in the event we *do* implicitly perform inference, there is some hedging. In this sense, the best uncertainty visualisation you can get without specifying a signal you want to convey is visualising the data itself. The census dot map's addition of interactivity also allows users to zoom in and see the details that caused "uncertainty" in the form of inconsistent colours at lower resolutions when they were zoomed out. This does not mean that visualising raw data instead of implementing sampling techniques is always a valid uncertainty visualisation that will prevent insignificant signal from getting through. @Buja2009 illustrated how groups that appear linearly separable in a linear discriminant analysis (LDA) visualisation of the data can actually be the result of a LDA performed on too many variables, something that was not clear from the visualisation until the line-up protocol was implemented. However it is simple but effective option that seems to be largely overlooked by the uncertainty visualisation community.

![Census dot map](censusdotmap.png){#fig-census}
